Bayesian Sparsification of $\cplx$-valued networks

### 2020-01-17 paper/paper.tex
read he at al. (2018), merged ell-0, and added citations

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
improvement methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods require additional considerations.
% the more explicitly zero parameters there is, the less computations is needed.

@@

data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\citep{hirose2003,fijulamin2009}
% [3] Akira Hirose, Complex-valued neural networks: theories and applications, vol. 5, World Scientific, 2003.
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

the complex version of backpropagation algorithm  was proposed because of
its richer representational capacity.
% Nevio Benvenuto and Francesco Piazza, “On the complex backpropagation algorithm,” IEEE Transactions on Signal Processing, vol. 40, no. 4, pp. 967–969, 1992.

@@

that interfaces the rest of
$\cplx$-valued
the network with complex numbers

that operates on the Fourier transform input speech, acoustic, of radio signals.

@@

Application of complex-valued networks is not new, \citep{citation_needed}. More recently
they were applied to music transcription and speech recognition \citep{trabelsi_deep_2017},
and \citep{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \citep{citation_needed},
acoustic and radio signal analysis \citep{citation_needed}, and quantum systems modelling
\citep{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.
%%% <<<< (cplx)

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
whihc updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Under suitable modelling and prior assumptions, Bayesian methods
can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \citep{he_amc:_2018}.
\todo{read (He et al., 2018)}
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch? yes}

Bayesian inference is a very rich field of research with a large body of knowledge and techniques
that enable practical quantification of predictive, parameter, or model uncertainty and its
propagation into solutions of the problems of interest.

The latter approach introduces stochastic masks one the parameters, which have hard-concrete
mask distribution. It is closely related to gumbel-softmax reparameterization trick,
\citep{kingma_auto-encoding_2014,jang_categorical_2017}: the mask is a stretched sigmoid,
clamped to $[0,1]$, that takes random input shifted relative to a learnable location
parameter, which indicates the degree of relevance. The penalty in the objective is a
smooth approximation to $\ell_0$ norm, i.e. the expected number of nonzero parameters.

% our contribution
Inspired by variational sparsification we develop bayesian dropout for $\cplx$-valued
networks.

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate and compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes in sec.~\ref{sec:discussion_and_related_research}.

@@

If we assume a mean field $\cplx$-Gaussian approximate posteriror for $
  W \in \cplx^{n\times m}
$, then
\begin{equation}  \label{eq:c-gauss-vi-general}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b_i,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
Hence \eqref{eq:cplx-gauss-trick} folows.

>>>>

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research.

@@

data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\citep{hirose2003,fijulamin2009}
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

the complex version of backpropagation algorithm  was proposed because of
its richer representational capacity.
% Nevio Benvenuto and Francesco Piazza, “On the complex backpropagation algorithm,” IEEE Transactions on Signal Processing, vol. 40, no. 4, pp. 967–969, 1992.
that operates on the Fourier transform input speech, acoustic, of radio signals.

Application of complex-valued networks is not new, \citep{citation_needed}. More recently
they were applied to music transcription and speech recognition \citep{trabelsi_deep_2017},
and \citep{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \citep{citation_needed},
acoustic and radio signal analysis \citep{citation_needed}, and quantum systems modelling
\citep{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.

@@

@@

%%% <<<< (cplx)

% motivation and other methods
Computational efficiency improvement and model compression methods such as quantization
\citep{uhlich_differentiable_2019}, integer-based arithmetic \citep{lin_fixed_2016,chen_fxpnet_2017},
and knowledge distillation \citep{hinton_distilling_2015} appear to be directly applicable
to $\cplx$-valued networks.
%
Certain model sparsification methods, e.g. sparsity inducing regularizers, and magnitude
pruning \citep{zuo_compression_2019}, which reactivates zeroed parameters based on gradients,
can be adapted directly to $\cplx$-valued models. Other methods require additional considerations.
% the more explicitly zero parameters there is, the less computations is needed.

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
have shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated multitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Bayesian inference is a very rich field of research with a large body of research and
techniques that enable practical quantification of predictive, parameter, or model
uncertainty and its propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Under suitable modelling and prior assumptions, Bayesian methods
can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$, \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that despite dropout showing more variable
performance, than magnitude pruning, it nevertheless induces non-uniform sparsity throughout
the model, which is essential for superior compression \citep{he_amc:_2018}.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational sparsification and stochastic sparsification techniques we
develop bayesian dropout for $\cplx$-valued networks.
%
The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate and compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes in sec.~\ref{sec:discussion_and_related_research}.

@@

If we assume a mean field $\cplx$-Gaussian approximate posteriror for $
  W \in \cplx^{n\times m}
$, then
\begin{equation}  \label{eq:c-gauss-vi-general-vec}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick-appendix}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b_i,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
Hence \eqref{eq:cplx-gauss-trick-appendix} folows.
